---
title: "Homework_1_StatMod"
name: Nathalie Guibert
date: "2023-10-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/whatt/iCloudDrive/UZH/Master/HS23/Polito/Statistical Modelling/Homeworks/Homework 1")

library(tidyverse)
library(fixest) 
library(lme4) 
library(sjPlot) 
library(texreg) 
library(lmtest) 
library(sandwich)
library(haven)
library(car)
library(stats)

ESS10 <- read_dta("ESS10.dta")
```

Data Cleaning and Filtering

```{r}
swiss_data <- filter(ESS10, cntry == "CH")

swiss_model <- subset(swiss_data, select = c(cntry, atchctr, prtvthch, prtdgcl, imdfetn, agea, gndr, pspwght))


# agea = Age of respondent, calculated  
summary(swiss_model$agea)
swiss_model <- subset(swiss_model, agea >= 18 )


# gndr = Gender 
summary(swiss_model$gndr)
swiss_model$gndr <- car::recode(swiss_model$gndr, "1=1; 2=2; else=NA")


# atchctr = How emotionally attached to [country] 
summary(swiss_model$atchctr)
swiss_model <- subset(swiss_model, atchctr <= 10)


# prtdgcl = How close to party 
summary(swiss_model$prtdgcl)
swiss_model <- subset(swiss_model, prtdgcl <= 4)


# imdfetn = Allow many/few immigrants of different 
# race/ethnic group from majority
summary(swiss_model$imdfetn)
swiss_model <- subset(swiss_model, imdfetn <= 4)


# prtvthch = Party voted for in last national election CH 
summary(swiss_model$prtvthch)

swiss_model <- swiss_model %>% 
  mutate(prtvthch = 
           case_when( 
                    prtvthch == 1~"SVP",
                    prtvthch == 2~"SP",
                    prtvthch == 3~"FDP",
                    prtvthch == 4~"CVP",
                    prtvthch == 5~"Gruene",
                    prtvthch == 6~"GLP",
                    prtvthch == 7~"BDP",
                    prtvthch == 8~"EVP",
                    prtvthch > 17~NA_character_,
                    TRUE ~ "other"
                  
                    )
        ) 




# Renaming of variables
swiss_model <- plyr::rename(swiss_model, replace = c(gndr = "Gender", agea = "Age", prtvthch = "Party_Vote_CH", atchctr = "Emot_Attach", prtdgcl = "Closeness_Party", imdfetn = "Immig_Openness"), warn_missing = FALSE)

```

1. Linear Regression Model

``` {r}

linreg_1 = lm(Emot_Attach ~ Party_Vote_CH + Age + Gender, data = swiss_model, weights = pspwght)
summary(linreg_1)

texreg(linreg_1)
```

2. Statistical Difference between SP & CVP

Solution: From the output of the linear hypothesis test, we cannot see a 
significant difference between CVP and SP concerning the emotional attachment 
to Switzerland as p>0.05 (0.6406).


``` {r}
linearHypothesis(linreg_1, "Party_Vote_CHCVP-Party_Vote_CHSP")
```

3. Influential data points 

We find 10 outliers in our model. 60 data points are influential, as well. 2 of
the outliers are influential.

Outliers
```{r}
sum(abs(rstudent(linreg_1)) > 3)
```

Hat values

```{r}
hat <- hatvalues(linreg_1)
sum(hat> 2 * mean(hat))
```

Influential Outliers

```{r}
sum(abs(rstudent(linreg_1)) > 3 & hat > 2 * mean(hat))
```

4. Simulate discrete changes

Solution: The discrete changes remain the same over age and gender.

```{r}
# selected party: SP

num <- 0

dc_df <- data.frame(dc=NA, age=18:80, lower=NA, upper=NA)

for(age in 18:80) {
  
  set.seed(21)
  num <- num + 1
  
  beta = MASS::mvrnorm(1000, coef(linreg_1), 
                         vcov(linreg_1))
  # HAS TO BE CHANGED IF WE CHANGE THE CATEGORIES OF THE PARTY VOTE ABOVE
  
  # MALE
  values_male = c(1, 0, 0, 0, 0, 0, 0, 1,0, age, 1)
  
  yhat_male =  beta %*% values_male
  
  
  # FEMALE
  
  values_female = c(1, 0, 0, 0, 0, 0, 0, 1,0, age, 2)
  
  yhat_female =  beta %*% values_female
  
  
  dc <- yhat_female - yhat_male
  
  
  dc_df[num,1] <- mean(dc)
  
  dc_df[num,3:4] <- quantile(dc, probs= c(0.025, 0.975))
  
}

ggplot(dc_df, aes(x=age, y=dc)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2)+
  geom_line(size= 0.5)+
  theme_minimal()+
  labs(y="Discrete Change", title = "Discrete change of gender, depending on 
       age for SP")

``` 


5. Estimate model again with variable closeness to party as IV.

Solution: The first model without the interaction seems Slightly better, 
because of higher R^2 values. Therefore, the first model explains more 
variance in the dependent variable than the second model with the interaction.


```{r}
summary(swiss_model$Closeness_Party)

swiss_model$Closeness_Party <- 
  car::recode(swiss_model$Closeness_Party, 
              "1=1; 2=2; 3:4=3; else=NA")


linreg_2 = lm(Emot_Attach ~ Party_Vote_CH + 
                Closeness_Party + Age + Gender, data = swiss_model, weights = pspwght)
summary(linreg_2)


# müssen Party vote und closeness einzeln 
# gelistet werden?
linreg_3 = lm(Emot_Attach ~ Party_Vote_CH*
                Closeness_Party + Age + Gender, data = swiss_model, weights = pspwght)
summary(linreg_3)


# Comparison

anova(linreg_2, linreg_3)


```

6. ICC & Multilevel Regression (Emotional attachment to 
country = DV)

```{r}
model_anova_hw = lmer(atchctr ~ 1 + (1 | cntry), data = ESS10, 
                   weights = pspwght)
summary(model_anova_hw)

# ICC
icc= 0.2579 / (0.2579 + 4.3731)
icc
```
Solution: Since the ICC value is 0.05568992, the country variable 
can barely explain emotional attachment, such that the individuals
are close to completely different.This means that roughly 5.57% comes 
from cross-country variation and the remains from individual variation.


7. Multilevel model 

Solution: Running the regression of the multilevel model, that people that 
allow some to none immigrants of different race/ethnic group in their country
are slightly more emotionally attached to their  country (0.2056467). This 
value is significant, given by its t-value 6.070, which exceeds the threshold 
of 1.96. Also with increasing age, there is also a slight positive effect on 
emotional attachment to one's country, which is also significant (t-value of 
34.859), however gender is not significant with a t-value below the threshold 
1.96 (0.481).

```{r}

ESS10$imdfetn_2 <- 
  car::recode(ESS10$imdfetn, 
              "1=1; 2:4=2; else=NA")
summary(ESS10$imdfetn_2)

mulreg_1 = lmer(atchctr ~ agea + gndr + imdfetn_2 + (1 | cntry), 
                data = ESS10, weights = pspwght)

summary(mulreg_1)
ranef(mulreg_1)
texreg(mulreg_1)


```
8. Random intercept plot

Solution: The following plot shows that Greece has the highest intercept and 
North Macedonia the lowest.

```{r}

plot_model(mulreg_1, type = "re", sort.est = "(Intercept)") + theme_bw()

```

9. Random slope w/ immigrant variable

Solution: We performed the likelihood ratio test and it was significant 
(6.61e-13) on the 0.000 level. This means that the random slope adds a
benefit to our model.

```{r}

model_rs = lmer(atchctr ~ agea + gndr + imdfetn_2 +(imdfetn_2 | cntry), data = ESS10,
               weights = pspwght)

model_rs

lrtest(mulreg_1, model_rs)
```

10. Fixed Effects

```{r}
# Fixed Effects Model
model_fixed_10 = feols(atchctr ~ agea + gndr + imdfetn_2 | cntry, 
                    data = ESS10, weights = ESS10$pspwght)

```

```{r}
# Fixed effects of Fixed Effects Model
fixef(model_fixed_10)
```

```{r}
summary(model_fixed_10)
```

```{r}
# Fixed effects of Random Effects Model
fixef(model_rs)
```

```{r}
# Random Effects of Random Effects Model
ranef(model_rs)
```